Quantifying Roughness in SAR Imagery with the Rényi Entropy

### 2025-03-29 Text/Articles/RENYI_ENTROPY_SAR_2025/R0-Renyi-entropy-sar.qmd
Update corrected comments

---
title: Quantifying Roughness in SAR Imagery with the Rényi Entropy
format:
  ieee-pdf:
    pdf-engine: pdflatex # xelatex
    keep-tex: true  
    #conference: true # comment this line to use journal
    #journaltype: conference # comment this line to use journal
    fig-cap-location: bottom # to crossref figure
    link-citations: true
    colorlinks: true
    linkcolor: black # equations
    citecolor: black # cites
    urlcolor: black
  #ieee-html: default

author:
  #- id: 
  - name: Janeth Alpala, Abraão D.&nbsp;C.&nbsp;Nascimento 
    affiliations:
      - name: Universidade Federal de Pernambuco
        department: Departamento de Estatística
        city: Recife 
        country: PE, Brazil
        postal-code: 50670-901
      #- name: Unknown affiliation
   # orcid: 
    email: janeth.alpala@ufpe.br, abraao@de.ufpe.br
    #url: 
    #membership: "Member, IEEE"
    #attributes:
    #  corresponding: true
    #photo: 
    #bio: |
    #  Use `IEEEbiography`  with figure as  option and
    #  the author name as the argument followed by the biography text.
  - name: Alejandro C.&nbsp;Frery
    affiliations:
      - name: Victoria University of Wellington
        department: School of Mathematics and Statistics
        city: Wellington 
        country: New Zealand
        postal-code: 6140
        orcid: "0000-0002-8002-5341"
        membership: "Fellow, IEEE"
      #- name: Unknown affiliation
    #orcid: 
    email: alejandro.frery@vuw.ac.nz
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Quantifying surface roughness in synthetic aperture radar (SAR) data is critical for accurate geophysical interpretation and remote sensing applications. 
  We propose a test statistic based on a non-parametric estimation of Rényi entropy to characterize surface roughness from SAR intensity data. 
  The statistic is refined using bootstrap to improve its stability, size and power. 
  This approach enhances roughness quantification by capturing scale-dependent variations and addressing data-driven uncertainty.
  Experimental results demonstrate the robustness of the proposed method in distinguishing roughness patterns, offering a statistically rigorous tool for SAR-based terrain analysis.
  <!-- Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems.  -->
  <!-- A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes.  -->
  <!-- In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed.  -->
  <!-- In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task.  -->
  <!-- In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy.  -->
  <!-- In particular, we use a non-parametric estimator for the Rényi entropy and investigate some of its properties.  -->
  <!-- As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories.  -->
  <!-- The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy. -->
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR, hypothesis tests, bootstrap]

@@

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   #- \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{mathtools}
  # - \usepackage[utf8]{inputenc}
   - \usepackage{hyperref}
   #- \hypersetup{draft} #Desactiva enlaces y referencias cruzadas
   - \usepackage{float}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{hyperref}
   - \hypersetup{
      colorlinks=true,
      linkcolor=black,
      filecolor=black,
      citecolor=black,
      urlcolor=black}

#bibliography: references.bib
bibliography: ../../Common/references.bib

@@

# Introduction
[S]{.IEEEPARstart}[ynthetic]{}
 Aperture Radar (SAR) technology has become essential for a wide range of applications&nbsp;[@Yu2023;@Akbarizadeh2012;@Mondini2021]. 
 It provides high-resolution data that is independent of sunlight and operates in a variety of weather conditions that facilitate global Earth monitoring&nbsp;[@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because this data is affected by speckle, a noise-like interference effect&nbsp;[@Argenti2013; @Choi2019]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

@@

Introduced by Alfréd Rényi in 1961&nbsp;[@renyi1961measures], this measure generalizes several well-known entropies, including Shannon's&nbsp;[@Ribeiro2021]. 
For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$, is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}
and
\begin{multline}\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr)=\ln\mu-\ln L +\ln(-1 - \alpha)\\
+ \frac{1}{1 - \lambda}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
      -\ln\Gamma(L)
   \bigr)\\
   +\ln\Gamma\bigl(\lambda(L - 1) + 1\bigr)
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)\\
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\end{multline}
These results are proven in the Appendix.
<!-- ACF Have you checked if it is possible to state the difference between \label{eq-HGI0} and \label{eq-HGammaSAR}? At the end of the day, that is what we want to assess. -->

@fig-plot presents the Rényi entropy of the $\mathcal{G}^0_I$ model as a function of $\mu$ for different values of $\alpha$. 
As $\alpha \to -\infty$, it approaches the Rényi entropy of the $\Gamma_{\text{SAR}}$ law, in line with the fact that the latter is a limiting case of the former.

```{r fig-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%",  fig.pos="hbt", fig.cap="$H_{\\lambda}(\\mathcal{G}^0_I)$ converges to the $H_{\\lambda}(\\Gamma_{\\text{SAR}})$ when $\\alpha\\to-\\infty$, with $L=8$ and $\\lambda=0.8$.", fig.width=4.5, fig.height=3.5}

@@

In order to estimate&nbsp;\eqref{E:entropy2} from a random sample, one may estimate the parameters that index the probability density function $f$, and integrate.
This approach heavily relies on the model assumptions and on the absence of outliers.
Alternatively, one may estimate directly $f$.
Such non-parametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics [@vasicek1976test; @Ebrahimi1994; @Wieczorkowski1999; @AlOmari2019]. 
Recently, Al-Labadi et al.&nbsp;[@AlLabadi2024] proposed a non-parametric estimator for Rényi entropy following this approach.

@@

We found that for $L > 1$ the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.
However, for $L = 1$, the optimal $\lambda$ tends to be higher and we chose $\lambda = 3$ to achieve good results. 
@fig-plotf illustrates the case for $L = 5$.

```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=7, fig.height=4.5}

@@

bias_plot <- ggplot(data, aes(x = Lambda, y = Bias)) +
  geom_line(color = "#00AFBB", linewidth = 1.0) +  
  geom_point(color = "#00AFBB", size = 3) +  
  geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$Bias), 
         #  label = "Bias", color = "#00AFBB", fontface = "italic", hjust = 1.0) +
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) +
  labs(x = expression(lambda), y = "Bias") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

mse_plot <- ggplot(data, aes(x = Lambda, y = MSE)) +
  geom_line(color = "#E69F00", linewidth = 1.0) +  
  geom_point(color = "#E69F00", size = 3) +
  #geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$MSE), 
           #label = "MSE", color = "#E69F00", fontface = "italic", hjust = 1.0) +
   scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) + 
  labs(x = expression(lambda), y = "MSE") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

@@

A Monte Carlo study with $1000$ samples from $\Gamma_{\text{SAR}}$, $\mu = 1$ and $L=5$, using $B = 200$ replications, confirms that the bootstrap-corrected estimator $\widetilde{H}_{\lambda}$, with $\lambda = 0.9$, outperforms the original $\widehat{H}_{\lambda}$ by reducing both bias and MSE for small sample sizes, as shown in @fig-Plot_bias_msef3.
<!-- ACF Remove the minor vertical grid (there is nothing between 49 and 81) -->

```{r Simulated_data_bias_B1, echo=FALSE, message=FALSE, cache = TRUE, autodep = TRUE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

Assume we have an estimator $\widetilde{H}_{\lambda}$ for the Rényi entropy of an arbitrary model. For testing \eqref{eq:hypothesis_test}, this estimator is expected to be close to the theoretical expression given in Equation \eqref{eq-HGammaSAR}. 
Since $L\geq1$ is known, we define the test statistic as follows:
\begin{multline}
\label{eq-test}
S(\bm{Z}; L) = \widetilde{H}_{\lambda} - \bigl\{\ln \widehat{\mu} - \ln L + \frac{1}{1-\lambda}
\bigl[-\lambda\,\ln\Gamma(L) \\ 
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  
- \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\bigr]\bigr\},
\end{multline}
where $\widehat{\mu}={n}^{-1}\sum_{i=1}^n Z_{i}$ is the sample mean.

When the null hypothesis holds, this test statistic is expected to be close to zero.
@fig-density_entropyR shows the empirical density of $S(\bm{Z}; L)$, based on $10^4$ replications of the $\Gamma_{\text{SAR}}$ model, with $\lambda=0.9$ and $L \in \{5,18\}$.
<!-- ACF 10^4 replications of samples of size?   -->

```{r Simulated_densityR, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = LegendOrder, linetype = LegendOrder, group = CurveOrder)) +
    geom_line(stat = "density", linewidth = 1.5) +
    scale_color_viridis(discrete = TRUE, option = "C", direction = -1, begin = 0.2, end = 0.7, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    scale_x_continuous(limits = x_limits, breaks = x_breaks) + 
    scale_y_continuous(limits = y_limits) +  
    labs(
        x = expression("Test Statistic" ~ S[widetilde(italic(H))[lambda]](italic(bold(Z))* ";" **phantom(" ")* italic(L))), 
        y = "Density"
    ) +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5, size = 16, margin = margin(b = 2)),
          axis.text = element_text(size = 16),     
        axis.title = element_text(size = 16),    
        legend.text = element_text(size = 16),   
        legend.title = element_text(size = 16)   
          )

@@

Vasicek&nbsp;[@vasicek1976test] proved that, for sufficiently large samples, $S(\bm{Z}; L)$ asymptotically follows a normal distribution:
\begin{equation*}
S(\bm{Z}; L) \overset{d}{\longrightarrow} \mathcal{N}(\mu_S,\,\sigma^{2}_S)\,,
\end{equation*}
where $d$ represents convergence in distribution. Here, $\mu_S  = \mathbb{E}[S(\bm{Z}; L)]$ and $\sigma^{2}_S = \text{Var}[S(\bm{Z}; L)]$ are the theoretical mean and variance of $S(\bm{Z}; L)$.

We define the standardized test statistic:
\begin{equation*}
\varepsilon = \frac{S(\bm{Z}; L) - \hat{\mu}_S}{\hat{\sigma}_S},
\end{equation*}
where $\hat{\mu}_S$ and $\hat{\sigma}_S$ are the empirical mean and standard deviation of $S(\bm{Z}; L)$, obtained by Monte Carlo simulations under the null hypothesis. 
By the central limit theorem, $\varepsilon$ asymptotically follows a standard normal distribution:
\begin{equation*}
\varepsilon \overset{d}{\longrightarrow} \mathcal{N}(0,1)\,.
\end{equation*}
Thus, the $p$-values are calculated as $2\Phi(-|\varepsilon|)$,
where $\Phi(\cdot)$ is the cumulative distribution function of the standard normal distribution.
@fig-density_entropyR_standardized0 shows smoothed histograms of standardized test statistics.

@@

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = LegendOrder, linetype = LegendOrder, group = CurveOrder)) +
    geom_line(stat = "density", linewidth = 1.3, alpha = 0.7) +
    scale_color_viridis(discrete = TRUE, option = "G", direction = -1, begin = 0.1, end = 0.7, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    scale_x_continuous(limits = x_limits, breaks = x_breaks) + 
    scale_y_continuous(limits = y_limits) +  
    labs(#x = expression("Standardized Test Statistic" ~ epsilon),
        x = expression("Standardized Test Statistic"), 
        y = "Density"
    ) +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5, size = 10, margin = margin(b = 2)),
          axis.text = element_text(size = 10),     
          axis.title = element_text(size = 10),    
          legend.text = element_text(size = 10),   
          legend.title = element_text(size = 10)   
    )

@@

    if (length(selected_columns) == 8) {
      combined_results <- combined_results[, selected_columns, drop = FALSE]
      colnames(combined_results) <- c("$\\bm{L}$", "$\\bm{n}$", 
                                      "$\\bm{1\\%}$", "$\\bm{5\\%}$", "$\\bm{10\\%}$", 
                                      "$\\bm{1\\%}$", "$\\bm{5\\%}$", "$\\bm{10\\%}$")
    }

>>>>

---
title: Quantifying Roughness in SAR Imagery with the Rényi Entropy
format:
  ieee-pdf:
    pdf-engine: pdflatex # xelatex
    keep-tex: true  
    classoption: lettersize
    tex-author-no-affiliation: true
    #conference: true # comment this line to use journal
    #journaltype: conference # comment this line to use journal
    fig-cap-location: bottom # to crossref figure
    link-citations: true
    colorlinks: true
    linkcolor: black # equations
    citecolor: black # cites
    urlcolor: black
  #ieee-html: default
author:
  - name: Janeth Alpala
    email: janeth.alpala@ufpe.br
    orcid: 0000-0002-0265-6236
  - name: Abraão D.&nbsp;C.&nbsp;Nascimento
    affiliations:
      - name: Universidade Federal de Pernambuco
        department: Departamento de Estatística
        city: Recife
        country: Brazil
        postal-code: 50670-901
    email: abraao@de.ufpe.br
    orcid: 0000-0003-2673-219X
  - name: Alejandro C.&nbsp;Frery
    affiliations:
      - name: Victoria University of Wellington
        department: School of Mathematics and Statistics
        city: Wellington
        country: New Zealand
        postal-code: 6140
    orcid: 0000-0002-8002-5341
    email: alejandro.frery@vuw.ac.nz
    membership: Fellow, IEEE
    attributes:
      corresponding: true
    note: |
      Janeth Alpala and Abraão D. C. Nascimento are with the Departamento de Estatística, Universidade Federal de Pernambuco, Recife, 50670-901 PE, Brazil (e-mails: janeth.alpala@ufpe.br, abraao@de.ufpe.br).

      Alejandro C. Frery is with the School of Mathematics and Statistics, Victoria University of Wellington, Wellington, 6140, New Zealand (e-mail: alejandro.frery@vuw.ac.nz). \emph{Corresponding author: Alejandro C. Frery.}

      Data and code are available at: <https://github.com/rjaneth/renyi-entropy> 
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Quantifying surface roughness in synthetic aperture radar (SAR) data is critical for accurate geophysical interpretation and remote sensing applications. 
  We propose a test statistic based on a non-parametric estimation of Rényi entropy to characterize surface roughness from SAR intensity data. 
  The statistic is refined using bootstrap to improve its stability, size and power. 
  This approach enhances roughness quantification by capturing scale-dependent variations and addressing data-driven uncertainty.
  Experimental results demonstrate the robustness of the proposed method in distinguishing roughness patterns, offering a statistically rigorous tool for SAR-based terrain analysis.
  <!-- Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems.  -->
  <!-- A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes.  -->
  <!-- In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed.  -->
  <!-- In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task.  -->
  <!-- In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy.  -->
  <!-- In particular, we use a non-parametric estimator for the Rényi entropy and investigate some of its properties.  -->
  <!-- As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories.  -->
  <!-- The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy. -->
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR, hypothesis tests, bootstrap]

@@

header-includes:
   #- \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{nccmath}
   - \usepackage{amssymb}
   - \usepackage{mathtools}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{hyperref}
   #- \hypersetup{draft} #Desactiva enlaces y referencias cruzadas
   - \usepackage{float}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{xcolor}
#bibliography: references.bib
bibliography: ../../Common/references.bib

@@

```
\renewcommand{\tablename}{TABLE}
# Introduction
[S]{.IEEEPARstart}[ynthetic]{}
 Aperture Radar (SAR) technology has become essential for a wide range of applications&nbsp;[@Yu2023;@Akbarizadeh2012;@Mondini2021]. 
 It provides high-resolution data that is independent of sunlight and operates in a variety of weather conditions that facilitate global Earth monitoring&nbsp;[@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because this data is affected by speckle, a noise-like interference effect&nbsp;[@Argenti2013; @Choi2019; @Baraha2023]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

@@

Introduced by Alfréd Rényi in 1961&nbsp;[@renyi1961measures], this measure generalizes several well-known entropies, including Shannon's&nbsp;[@Ribeiro2021]. 
For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$, is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using&nbsp;\eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and the $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}
and
\begin{multline}  
\label{eq-HGI0}  
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \ln(-1 - \alpha) \\  
+ \frac{1}{1 - \lambda} \Bigl[ \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) \\ 
+ \ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr) - \ln\Gamma\bigl(\lambda(L - \alpha)\bigr) \\
+ \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}
The derivation of the latter result is proven in the Appendix.
Note that the Rényi entropy of $\mathcal{G}^0_I$ can be expressed in terms of that of $\Gamma_{\mathrm{SAR}}$, with additional terms depending on the roughness parameter $\alpha$. As $\alpha \to -\infty$, these terms tend to zero, reducing&nbsp;\eqref{eq-HGI0} to&nbsp;\eqref{eq-HGammaSAR}.

@fig-plot illustrates this convergence by showing $H_\lambda(\mathcal{G}^0_I)$  as a function of $\mu$ for different values of $\alpha$. As $\alpha$ decreases, the curves approach $H_\lambda(\Gamma_{\text{SAR}})$ (solid black line). This confirms that $\Gamma_{\text{SAR}}$ is a limiting case of $\mathcal{G}^0_I$.
<!-- ACF Have you checked if it is possible to state the difference between \label{eq-HGI0} and \label{eq-HGammaSAR}? At the end of the day, that is what we want to assess. -->
<!-- Corrected -->
```{r fig-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%",  fig.pos="hbt", fig.cap="$H_{\\lambda}(\\mathcal{G}^0_I)$ converges to the $H_{\\lambda}(\\Gamma_{\\text{SAR}})$ when $\\alpha\\to-\\infty$, with $L=8$ and $\\lambda=0.8$.", fig.width=4.5, fig.height=3.5}

@@

In order to estimate&nbsp;\eqref{E:entropy2} from a random sample, one may estimate the parameters that index the probability density function $f$, and integrate.
This approach heavily relies on the model assumptions and on the absence of outliers.
Alternatively, one may estimate directly $f$.
Such non-parametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics&nbsp;[@vasicek1976test;@Bert1992; @Ebrahimi1994; @Wieczorkowski1999; @IbrahimAlOmari2014]. 
Recently, Al-Labadi et al.&nbsp;[@AlLabadi2024] proposed a non-parametric estimator for Rényi entropy following this approach.

@@

We found that for $L > 1$ the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.
However, for $L = 1$, the optimal $\lambda$ tends to be higher and we chose $\lambda = 3$ to achieve good results. 
@fig-plotf illustrates the case for $L = 5$.
```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=7, fig.height=4.5}

@@

bias_plot <- ggplot(data, aes(x = Lambda, y = Bias)) +
  geom_line(color = "#00AFBB", linewidth = 1.0) +  
  geom_point(color = "#00AFBB", size = 3) +  
  geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$Bias), 
         #  label = "Bias", color = "#00AFBB", fontface = "italic", hjust = 1.0) +
  #scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) +
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda, minor_breaks = NULL)+
  scale_y_continuous(minor_breaks = NULL)+
  labs(x = expression(lambda), y = "Bias") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

mse_plot <- ggplot(data, aes(x = Lambda, y = MSE)) +
  geom_line(color = "#E69F00", linewidth = 1.0) +  
  geom_point(color = "#E69F00", size = 3) +
  #geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$MSE), 
           #label = "MSE", color = "#E69F00", fontface = "italic", hjust = 1.0) +
   #scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) + 
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda, minor_breaks = NULL)+
  scale_y_continuous(minor_breaks = NULL)+
  labs(x = expression(lambda), y = "MSE") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

@@

A Monte Carlo study with $1000$ replications for each sample size $n \in \{9, 25, 49, 81, 121\}$ from the $\Gamma_{\text{SAR}}$ ($\mu=1, L=5$) confirms that for $\lambda=0.9$, the bootstrap-corrected estimator $\widetilde{H}_{\lambda}$ ($B=200$) reduces both bias and MSE compared to the original $\widehat{H}_{\lambda}$, with significant improvements for small sample sizes, as shown in&nbsp;@fig-Plot_bias_msef3.

<!-- ACF Remove the minor vertical grid (there is nothing between 49 and 81) -->
<!-- Corrected -->
```{r Simulated_data_bias_B1, echo=FALSE, message=FALSE, cache = TRUE, autodep = TRUE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

Assume we have an estimator $\widetilde{H}_{\lambda}$ for the Rényi entropy of an arbitrary model. For testing&nbsp;\eqref{eq:hypothesis_test}, this estimator is expected to be close to the theoretical expression given in Equation&nbsp;\eqref{eq-HGammaSAR}. 
Since $L\geq1$ is known, we define the test statistic as follows:
\begin{multline}
\label{eq-test}
S_{\widetilde{H}_{\lambda}}(\bm{Z}; L) = \widetilde{H}_{\lambda} - \bigl\{\ln \widehat{\mu} - \ln L + \frac{1}{1-\lambda}
\bigl[-\lambda\,\ln\Gamma(L) \\ 
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  
- \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\bigr]\bigr\},
\end{multline}
where $\widehat{\mu}={n}^{-1}\sum_{i=1}^n Z_{i}$ is the sample mean.
When the null hypothesis holds, this test statistic is expected to be close to zero.
@fig-density_entropyR shows the empirical density of $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, based on $10^4$ replications of the $\Gamma_{\text{SAR}}$ model for each sample size $n \in \{49,81, 121\}$ with $\lambda=0.9$ and $L \in \{5,18\}$.
<!-- ACF 10^4 replications of samples of size?   -->
<!-- Corrected -->
```{r Simulated_densityR, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = LegendOrder, linetype = LegendOrder, group = CurveOrder)) +
    geom_line(stat = "density", linewidth = 1.5) +
    scale_color_viridis(discrete = TRUE, option = "C", direction = -1, begin = 0.2, end = 0.7, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    scale_x_continuous(limits = x_limits, breaks = x_breaks, minor_breaks = NULL) +
    #scale_x_continuous(limits = x_limits, breaks = x_breaks) + 
    scale_y_continuous(limits = y_limits, minor_breaks = NULL)+
    #scale_y_continuous(limits = y_limits) +  
    labs(
        x = expression("Test Statistic" ~ S[widetilde(italic(H))[lambda]](italic(bold(Z))* ";" **phantom(" ")* italic(L))), 
        y = "Density"
    ) +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5, size = 16, margin = margin(b = 2)),
          axis.text = element_text(size = 16),     
        axis.title = element_text(size = 16),    
        legend.text = element_text(size = 16),   
        legend.title = element_text(size = 16)   
          )

@@

Vasicek&nbsp;[@vasicek1976test] proved that, for sufficiently large samples, $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$ asymptotically follows a normal distribution:
\begin{equation*}
S_{\widetilde{H}_{\lambda}}(\bm{Z}; L) \overset{d}{\longrightarrow} \mathcal{N}(\mu_S,\,\sigma^{2}_S)\,,
\end{equation*}
where $d$ represents convergence in distribution. Here, $\mu_S  = \mathbb{E}[S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)]$ and $\sigma^{2}_S = \text{Var}[S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)]$ are the theoretical mean and variance of $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$.

We define the standardized test statistic:
\begin{equation*}
\varepsilon = \frac{S_{\widetilde{H}_{\lambda}}(\bm{Z}; L) - \hat{\mu}_S}{\hat{\sigma}_S},
\end{equation*}
where $\hat{\mu}_S$ and $\hat{\sigma}_S$ are the empirical mean and standard deviation of $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, obtained by Monte Carlo simulations under the null hypothesis. 
By the central limit theorem, $\varepsilon$ asymptotically follows a standard normal distribution:
\begin{equation*}
\varepsilon \overset{d}{\longrightarrow} \mathcal{N}(0,1)\,.
\end{equation*}
Thus, the $p$-values are calculated as $2\Phi(-|\varepsilon|)$,
where $\Phi(\cdot)$ is the cumulative distribution function of the standard normal distribution.
@fig-density_entropyR_standardized0 shows smoothed histograms of standardized test statistics.

@@

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = LegendOrder, linetype = LegendOrder, group = CurveOrder)) +
    geom_line(stat = "density", linewidth = 1.3, alpha = 0.7) +
    scale_color_viridis(discrete = TRUE, option = "G", direction = -1, begin = 0.1, end = 0.7, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    #scale_x_continuous(limits = x_limits, breaks = x_breaks) + 
    #scale_y_continuous(limits = y_limits) +
    scale_x_continuous(limits = x_limits, breaks = x_breaks, minor_breaks = NULL) +
scale_y_continuous(limits = y_limits, minor_breaks = NULL)+
    labs(#x = expression("Standardized Test Statistic" ~ epsilon),
        x = expression("Standardized Test Statistic"), 
        y = "Density"
    ) +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5, size = 10, margin = margin(b = 2)),
          axis.text = element_text(size = 10),     
          axis.title = element_text(size = 10),    
          legend.text = element_text(size = 10),   
          legend.title = element_text(size = 10)   
    )

@@

    if (length(selected_columns) == 8) {
      combined_results <- combined_results[, selected_columns, drop = FALSE]
      colnames(combined_results) <- c("$\\bm{L}$", "$\\bm{n}$", 
  "$\\hphantom{00}\\SI{1}{\\percent}$", 
  "$\\hphantom{00}\\SI{5}{\\percent}$", 
  "$\\hphantom{00}\\SI{10}{\\percent}$", 
  "$\\hphantom{00}\\SI{1}{\\percent}$", 
  "$\\hphantom{00}\\SI{5}{\\percent}$", 
  "$\\hphantom{00}\\SI{10}{\\percent}$")
    }
